{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-26 thesis/education.tex
mechanics proofreading - function names, footnote placement, figure optmz, syn hi

\begin{itemize}
	\item Finds the declaration of {\tt ready\_list}, the scheduler runqueue declared by the basecode,
		%and detects if the student has modified to be an array of lists rather than a single one.
		\revisionminor{and detects if the student has replaced the default definition
		with a different name and/or data structure.}
		If so,
		%defines the length of that array in a macro to be used by {\tt is\_runqueue()}
		\revisionminor{emits macros to configure {\tt is\_runqueue()}'s behaviour to handle
		certain common alternate approaches}
		(part of the patch described below).
		Either way defines a function {\tt get\_rq\_addr()} to return the address of the %(first)
		list.
	\item Changes the basecode's definition of {\tt TIME\_SLICE} from 4 to 1 (units of timer ticks)
		so Landslide's timer injection will properly drive the context switcher.
	\item Inserts {\tt tell\_landslide\_forking()} into {\tt thread.c}
		(using {\tt sed} rather than the patch, described below,
		because it must go in a function which students have to implement,
		which is likely to disturb the context and make a patch fail).
	\item Adds the new {\tt priority-donate-multiple} test.
	\item Applies the {\tt annotate-pintos.patch} patch to the imported student implementation, which:
	\begin{itemize}
		\llitem Adds {\tt tell\_landslide\_thread\_on\_rq()}
			and {\tt tell\_landslide\_thread\_off\_rq()}
			annotations
			to {\tt list\_insert()} and {\tt list\_remove()} respectively
			(in {\tt lib/kernel/\allowbreak{}list.c}, which the students don't modify),
			which
			check whether the argument list
			is the scheduler runqueue
			using a helper function {\tt is\_runqueue},
			which in turn uses {\tt get\_rq\_addr()} and {\tt READY\_LIST\_LENGTH} described above.
		\llitem Modifies the existing {\tt priority-sema} and {\tt alarm-simultaneous} tests to be more Landslide-friendly.
		\llitem Inserts the {\tt tell\_landslide\_sched\_init\_done()},
			{\tt tell\_landslide\_vanishing()},
			and {\tt tell\_landslide\_thread\_switch()}
			annotations in the appropriate places
			(which the students generally do not modify).
	\end{itemize}
	\item Detects if the student has renamed the {\tt elem} field of the TCB struct,
		and if so renames its use in {\tt is\_runqueue()} (described above) correspondingly.
	\item Detects if the student has renamed the {\tt cur} (currently running thread) variable
		in the context switcher, and if so renames it back.
\end{itemize}

@@

\begin{itemize}
	\item {\tt priority-sema}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads to wait on a semaphore and signals them.
		Replaces threads with different priorities
		(originally chosen to produce deterministic output which the test checked for)
		with threads of the same priority.
	\item {\tt alarm-simultaneous}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads which each invoke {\tt timer\_sleep()} for a different amount of time.
		Number of (threads,iterations) reduced from (3,5) to (2,1).
	\item {\tt wait-simple}:
		Unmodified from the basecode's version, for {\em userprog}.
		Userspace process {\tt exec()}s a child process, which immediately exits, and {\tt wait()}s on it.
	\item {\tt wait-twice}:
		Unmodified from the basecode's version, for {\em userprog}.
		Slightly more complicated version of {\tt wait-simple},
		intended to expose failure-path bugs if the former finds no easier ones.
	\item {\tt priority-donate-multiple}:
		Written by Kevin Zhao, TA at U. Chicago, for {\em threads}.
		Tests for a priority donation race during {\tt lock\_release()}
		in which a thread holding a lock can accidentally keep a contending thread's donated priority
		after finishing releasing it.
\end{itemize}
\vspace{1em}

@@

\begin{enumerate}
	\item {\tt priority-donate-multiple} found \revisionminor{two}
		deterministic bugs and \revisionminor{one} concurrency bug in {\tt threads} projects.
		It found the targeted priority-leak bug
		described in \cref{sec:education-pintos-tests} in one group,
		and exposed a {\tt NULL} pointer dereference in another
		when attempting to donate priority to an already-exited thread.
		The latter bug was observed deterministically.
		A third group neglected to implement priority donation during {\tt lock\_release()} at all,
		which was found deterministically.
	\item {\tt wait-simple} found \revisionminor{one} deterministic bug
		and \revisionminor{three} concurrency bugs in {\tt userprog} projects.
		It found a deadlock in two groups' implementations,
		% for one of these groups 2 traces were emitted, same stack trace (i think) just w diff pps
		wherein {\tt process\_\allowbreak{}wait()}'s synchronization assumed
		it would always be called before {\tt process\_exit()}.
		In both cases the former's {\tt cond\_wait()} call would block forever if reordered after the latter.
		This test also found several use-after-frees for one of those two groups,
		whose {\tt process\_exit()} didn't guard against a parent process exiting and deallocating its memory first;
		this root cause manifested in heap errors in \revisionminor{three} separate locations.
		Lastly, the heap checking alone found a deterministic use-after-free in a third group's {\tt exec()}.
\end{enumerate}

@@

{\bf Statistical significance.}
\Cref{tab:table-of-ze-studence} presents more detailed statistics from these six distributions.
To assess whether the differences are significant,
% discrete == I used "version 2" of the AD metric that R outputs.
I use the $k$-sample discrete Anderson-Darling test \cite{anderson1952,anderson-darling}
provided by the R statistical programming environment \cite{r-lang,r-ksamples},
comparing each control group to its respective Landslide users group.
Anderson-Darling uses a weighted sum-of-squares distance metric to compare two empirical distribution functions (EDFs),
testing if their samples are likely to arise from the same underlying, though unspecified, distribution.%
\footnote{
I choose Anderson-Darling over the simpler Kolmogorov-Smirnov,
which computes only the maximum instantaneous distance between EDFs,
for its better sensitivity both to repeated deviations and to tail differences \cite{beware-kolmogorov-smirnov}.
}
I deem the difference in grade distributions significant when $p < 0.05$%
% FIXME: is this legit?
% should you be using the combined/grouped AD test instead?
\footnote{I choose not to correct for multiplicity among these four comparsons \revisionminor{as} I do in the next section
because each tests a different hypothesis.
},
although as in all $p$-value calculations,
this rejects only the null hypothesis,
not alternative hypotheses,
such as ``more skilled students were more likely to sign up to begin with''.
%
Ultimately, only the same-semester comparisons,
\revisionminor{which do not address that alternative hypothesis,
were significant,
while the cross-semester comparisons,
which attempted to do so,
failed to confidently reject even the null hypothesis.}
However, the distribution difference being smaller in P3 than in P2
further suggests, albeit informally,
that the larger impact in P2 is not attributable {\em only} to selection bias,
or else a similar difference should have been observed in the P3 distributions.
Another possible explanation could simply be that the P3 grading critera result in less grade variance overall,
but this pattern is at least consistent with the ``Landslide helps students submit better P2s'' hypothesis.

@@

\begin{enumerate}
	\item Requested for preemption traces to be more clear about the meaning of each stack trace in each table cell,
		and complained of inaccurate line numbers
		[likely referring to how the current behaviour indicates the line of code {\em after} a function call,
		corresponding to the {\tt call} instruction's pushed return address,
		rather than the function call itself]. [This answer from CMU; see above.]
	\item Including a manual or tutorial would be helpful [presumably beyond the user guide's instructions, such as recapping the procedure shown in the lecture demo which wasn't written down anywhere].
	\item Don't print warnings about line length exceeding 80 characters [inconsistency between 15-410 and \psuos compilation options].
	\item ``It takes too long. But I guess that's impossible to fix.'' [Well, it's an open research problem to fix!]
	\item Preemption traces should explicitly indicate where in the interleaving the bug occurred.
		[Root cause identification is its own research area, but more detail is certainly possible.]
	\item Improve explanation of data races [in the user guide, perhaps].
	\item Happy with it as-is [\revisionminor{reported by three} students]
		% wow, latex -- if you put [] outside of the revision, everything is awful forever
	\item \revisionminor{[Seven non-respondents]}
\end{enumerate}

>>>>

\begin{itemize}
	\item Finds the declaration of {\tt ready\_list}, the scheduler runqueue declared by the basecode,
		%and detects if the student has modified to be an array of lists rather than a single one.
		\revisionminor{and detects if the student has replaced the default definition
		with a different name and/or data structure.}
		If so,
		%defines the length of that array in a macro to be used by {\tt is\_runqueue()}
		\revisionminor{emits macros to configure {\tt is\_runqueue()} to handle
		certain common alternate approaches}
		(part of the patch described below).
		Either way,
		defines a function {\tt get\_rq\_addr()} to return the address of the %(first)
		list.
	\item Changes the basecode's definition of {\tt TIME\_SLICE} from 4 to 1 (units of timer ticks)
		so Landslide's timer injection will properly drive the context switcher.
	\item Inserts {\tt tell\_landslide\_forking()} into {\tt thread.c}
		(using {\tt sed} rather than the patch, described below,
		because it must go in a function which students have to implement,
		which is likely to disturb the context and make a patch fail).
	\item Adds the new {\tt priority-donate-multiple} test.
	\item Applies the {\tt annotate-pintos.patch} patch to the imported student implementation, which:
	\begin{itemize}
		\llitem Adds {\tt tell\_landslide\_thread\_on\_rq()}
			and {\tt tell\_landslide\_thread\_off\_rq()}
			annotations
			to {\tt list\_insert()} and {\tt list\_remove()} respectively
			(in {\tt lib/kernel/\allowbreak{}list.c}, which the students don't modify),
			which
			check whether the argument list
			is the scheduler runqueue
			using a helper function {\tt is\_runqueue},
			which in turn uses {\tt get\_rq\_addr()} and {\tt READY\_LIST\_LENGTH} described above.
		\llitem Modifies the existing {\tt priority-sema} and {\tt alarm-simultaneous} tests to be more Landslide-friendly.
		\llitem Inserts the {\tt tell\_landslide\_sched\_init\_done()},
			{\tt tell\_landslide\_vanishing()},
			and {\tt tell\_landslide\_thread\_switch()}
			annotations in the appropriate places
			(which the students generally do not modify).
	\end{itemize}
	\item Detects if the student has renamed the {\tt elem} field of the TCB struct,
		and if so renames its use in {\tt is\_runqueue()} (described above) correspondingly.
	\item Detects if the student has renamed the {\tt cur} (currently running thread) variable
		in the context switcher, and if so renames it back.
\end{itemize}

@@

\begin{itemize}
	\item {\tt priority-sema}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads to wait on a semaphore and signals them.
		Replaces threads with different priorities
		(originally chosen to produce deterministic output which the test checked for)
		with threads of the same priority.
	\item {\tt alarm-simultaneous}:
		Modified to be Landslide-friendly from the basecode, for {\em threads}.
		Creates two child threads which each invoke {\tt timer\_sleep()} for a different amount of time.
		Number of (threads,iterations) reduced from (3,5) to (2,1).
	\item {\tt wait-simple}:
		Unmodified from the basecode's version, for {\em userprog}.
		Userspace process {\tt exec}s a child process, which immediately exits, and {\tt wait}s on it.
	\item {\tt wait-twice}:
		Unmodified from the basecode's version, for {\em userprog}.
		Slightly more complicated version of {\tt wait-simple},
		intended to expose failure-path bugs if the former finds no easier ones.
	\item {\tt priority-donate-multiple}:
		Written by Kevin Zhao, TA at U. Chicago, for {\em threads}.
		Tests for a priority donation race during {\tt lock\_release()}
		in which a thread holding a lock can accidentally keep a contending thread's donated priority
		after finishing releasing it.
\end{itemize}
\vspace{1em}

@@

\begin{enumerate}
	\item {\tt priority-donate-multiple} found \revisionminor{two}
		deterministic bugs and \revisionminor{one} concurrency bug in {\tt threads} projects.
		It found the targeted priority-leak bug
		described in \cref{sec:education-pintos-tests} in one group,
		and exposed a {\tt NULL} pointer dereference in another
		when attempting to donate priority to an already-exited thread.
		The latter bug was observed deterministically.
		A third group neglected to implement priority donation during {\tt lock\_release()} at all,
		which was found deterministically.
	\item {\tt wait-simple} found \revisionminor{one} deterministic bug
		and \revisionminor{three} concurrency bugs in {\tt userprog} projects.
		It found a deadlock in two groups' implementations,
		% for one of these groups 2 traces were emitted, same stack trace (i think) just w diff pps
		wherein {\tt process\_\allowbreak{}wait()}'s synchronization assumed
		it would always be called before {\tt process\_exit()}.
		In both cases the former's {\tt cond\_wait()} call would block forever if reordered after the latter.
		This test also found several use-after-frees for one of those two groups,
		whose {\tt process\_exit()} didn't guard against a parent process exiting and deallocating its memory first;
		this root cause manifested in heap errors in \revisionminor{three} separate locations.
		Lastly, the heap checking alone found a deterministic use-after-free in a third group's {\tt process\_execute()}.
\end{enumerate}

@@

{\bf Statistical significance.}
\Cref{tab:table-of-ze-studence} presents more detailed statistics from these six distributions.
To assess whether the differences are significant,
% discrete == I used "version 2" of the AD metric that R outputs.
I use the $k$-sample discrete Anderson-Darling test \cite{anderson1952,anderson-darling}
provided by the R statistical programming environment \cite{r-lang,r-ksamples},
comparing each control group to its respective Landslide users group.
Anderson-Darling uses a weighted sum-of-squares distance metric to compare two empirical distribution functions (EDFs),
testing if their samples are likely to arise from the same underlying, though unspecified, distribution.%
\footnote{
I choose Anderson-Darling over the simpler Kolmogorov-Smirnov,
which computes only the maximum instantaneous distance between EDFs,
for its better sensitivity both to repeated deviations and to tail differences \cite{beware-kolmogorov-smirnov}.
}
I deem the difference in grade distributions significant when $p < 0.05$,%
% FIXME: is this legit?
% should you be using the combined/grouped AD test instead?
\footnote{I choose not to correct for multiplicity among these four comparsons \revisionminor{as} I do in the next section
because each tests a different hypothesis.
}
although as in all $p$-value calculations,
this rejects only the null hypothesis,
not alternative hypotheses,
such as ``more skilled students were more likely to sign up to begin with''.
%
Ultimately, only the same-semester comparisons,
\revisionminor{which do not address that alternative hypothesis,
were significant,
while the cross-semester comparisons,
which attempted to do so,
failed to confidently reject even the null hypothesis.}
However, the distribution difference being smaller in P3 than in P2
further suggests, albeit informally,
that the larger impact in P2 is not attributable {\em only} to selection bias,
or else a similar difference should have been observed in the P3 distributions.
Another possible explanation could simply be that the P3 grading critera result in less grade variance overall,
but this pattern is at least consistent with the ``Landslide helps students submit better P2s'' hypothesis.

@@

\begin{enumerate}
	\item Requested for preemption traces to be more clear about the meaning of each stack trace in each table cell,
		and complained of inaccurate line numbers
		[likely referring to how the current behaviour indicates the line of code {\em after} a function call,
		corresponding to the {\tt call} instruction's pushed return address,
		rather than the function call itself]. [This answer from CMU; see above.]
	\item Including a manual or tutorial would be helpful [presumably beyond the user guide's instructions, such as recapping the procedure shown in the lecture demo which wasn't written down anywhere].
	\item Don't print warnings about line length exceeding 80 characters [inconsistency between 15-410 and \psuos compilation options].
	\item ``It takes too long. But I guess that's impossible to fix.'' [Well, it's an open research problem to fix!]
	\item Preemption traces should explicitly indicate where in the interleaving the bug occurred.
		[Root cause identification is its own research area, but more detail is certainly possible.]
	\item Improve explanation of data races [in the user guide, perhaps].
	\item Happy with it as-is [\revisionminor{reported by three} students].
		% wow, latex -- if you put [] outside of the revision, everything is awful forever
	\item \revisionminor{[Seven non-respondents.]}
\end{enumerate}

### 2018-12-26 thesis/future.tex
mechanics proofreading - function names, footnote placement, figure optmz, syn hi

As important as Quicksand's convergence theorem (\cref{sec:quicksand-soundness}),
the HTM equivalence proofs (\cref{sec:tm-design}),
and other soundness results from prior work are,
they fall short of fully end-to-end formal verification
in terms of trusting the output of stateless model checkers in practice.
In addition to the issue of test cases,
%discussed above,
there is also currently no guarantee that Landslide's concrete implementation
matches the theoretical properties.
My personal faith in Landslide's implementation comes from years of empirical observation:
inspecting small state spaces
(such as \Cref{fig:retry-reduction})
by hand
to check that preemption points show up in the right places
and that DPOR prunes the expected equivalences;
as well as checking that the relationships between larger state spaces satisfy expected invariants,
such as when testing two sets of preemption points, one a subset of the other,
or testing the same but one with an additional reduction applied,
% this first "the" realllly helps disambiguate the sentence
the one's resulting state space must be smaller than the other's.
Nevertheless, discrepancies may still lurk undiscovered:
the failure injection bug described in \cref{sec:tm-eval-exp-setup}
took until implementing retry set reduction to discover,
well after publishing the first experimental results under it in \cite{sigbovik-htm}.
As the concurrency model becomes more complex,
more opportunities arise for the implementation to deviate from what the soundness theorem actually proves.
Kernels like seL4 \cite{sel4} and CertiKOS \cite{certikos}
and concurrency-aware compilers like CompCertTSO \cite{compcerttso},
all formally verified against external specifications,
provide important pioneering steps in this direction.%
\footnote{Of course, being verified themselves, these projects have no need for a verified Landslide,
but checking unverified (e.g., student) code with such would still be an improvement.}

>>>>

As important as Quicksand's convergence theorem (\cref{sec:quicksand-soundness}),
the HTM equivalence proofs (\cref{sec:tm-design}),
and other soundness results from prior work are,
they fall short of fully end-to-end formal verification
in terms of trusting the output of stateless model checkers in practice.
In addition to the issue of test cases,
%discussed above,
there is also currently no guarantee that Landslide's concrete implementation
matches the theoretical properties.
My personal faith in Landslide's implementation comes from years of empirical observation:
inspecting small state spaces
(such as \Cref{fig:retry-reduction})
by hand
to check that preemption points show up in the right places
and that DPOR prunes the expected equivalences;
as well as checking that the relationships between larger state spaces satisfy expected invariants,
such as when testing two sets of preemption points, one a subset of the other,
or testing the same but one with an additional reduction applied,
% this first "the" realllly helps disambiguate the sentence
the one's resulting state space must be smaller than the other's.
Nevertheless, discrepancies may still lurk undiscovered:
the failure injection bug described in \cref{sec:tm-eval-exp-setup}
took until implementing retry set reduction to discover,
well after publishing the first experimental results in \cite{sigbovik-htm}.
As the concurrency model becomes more complex,
more opportunities arise for the implementation to deviate from what the soundness theorem actually proves.
Kernels like seL4 \cite{sel4} and CertiKOS \cite{certikos}
and concurrency-aware compilers like CompCertTSO \cite{compcerttso},
all formally verified against external specifications,
provide important pioneering steps in this direction.%
\footnote{Of course, being verified themselves, these projects have no need for a verified Landslide,
but checking unverified (e.g., student) code with such would still be an improvement.}

### 2018-12-26 thesis/quicksand.tex
mechanics proofreading - function names, footnote placement, figure optmz, syn hi

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{p{0.9\textwidth}}
			\begin{center}
			\begin{tabular}{rlcrl}
				\multicolumn{5}{c}{Initially {\tt int count = \const{0}; mutex\_t m;}} \\
				\\
				& {\bf Thread 1} & & & {\bf Thread 2} \\
				\hline
				11 & \texttt{\hilight{darkorange}{mutex\_lock}(\&m);} & &
						21 & \texttt{atomic\_xadd(\&count, \const{1});} \\
				12 & \texttt{count++;}                                & &
						22 & \texttt{\hilight{olivegreen}{yield}();} \\
				13 & \texttt{\hilight{darkblue}{mutex\_unlock}(\&m);} & &
						23 & \texttt{atomic\_xadd(\&count, \const{1});} \\
				14 & \texttt{assert(count >= \const{1});}             & &
						24 & \texttt{assert(count >= \const{2});} \\
			\end{tabular}
			\end{center}
			\\
			\begin{center}
			(a) Example program with synchronization API calls highlighted.
			\end{center}
			\\
			\begin{center}
			\begin{tabular}{llllc}
				& {\bf Thread 1} & & {\bf Thread 2} & {\bf Value of {\tt count}} \\
				\hline
				11 & \texttt{\hilight{darkorange}{mutex\_lock}(\&m);} & & & 0 \\
				12a & \texttt{int tmp = count;} & & & 0 \\
							& & 21 & \texttt{atomic\_xadd(\&count, \const{1});} & 1 \\
							& & 22 & \texttt{\hilight{olivegreen}{yield}();} & 1 \\
							& & 23 & \texttt{atomic\_xadd(\&count, \const{1});} & 2 \\
				12b & \texttt{count = tmp + 1;} & & & 1 \\
				13 & \texttt{\hilight{darkblue}{mutex\_unlock}(\&m);} & & & 1 \\
				14 & \texttt{assert(count >= \const{1});} & & & 1 \\
							& & 24 & \texttt{\hilight{assertfail}{assert(count >= 2);}} & 1 \\
			\end{tabular}
			\end{center}
			\\
			%(b) Interleaving of (a) in which Thread 2's assertion fails,
			%as the single increment of Thread 1 overwrites both those of Thread 2.
			(b) Buggy interleaving of (a) in which
			the single increment of Thread 1 overwrites both those of Thread 2.
			Note the data-race preemption at 12a-12b.
		\end{tabular}
	\end{center}
	\caption[Example bug requiring data-race preemption points to expose.]
	{Example bug requiring data-race preemption points to expose.
	Because the two threads use different modes of synchronization to protect their respective accesses to {\tt count},
	preempting on synchronization calls alone is insufficient to expose the bug.
	Rather, Thread 1 must be preempted between its non-atomic load and store of {\tt count}.
	}
	\label{fig:pps-example}
\end{figure}

@@

On 229 tests, SSS-MC-ICB timed out after 10 hours with no bugs found. % thesis note - pretty sure this *isn't* preempt-everywhere
Among these tests, QS-Limited-HB % thesis note -- i think that's what past-me meant. they just wrote, 'quicksand' but.
found bugs in 37.
The other 192 represent cases where neither Quicksand nor ICB were able to provide a conclusive result either way.
\footnote{Thesis note: SSS-MC-Shared-Mem was added
%in a subsequent revision to \cite{quicksand}
%from when this analysis was conducted
subsequently to this analysis's publication in \cite{quicksand},
at which time SSS-MC-ICB was the best-performing approach among control experiments.
Nevertheless, mutual timeouts among all six testing approaches constituted roughly one third of the test suite.}
For these 192,
I show the number of state spaces Quicksand was able to complete in the
``Average tested subset SSes'' column.
% of \Cref{tab:drbugs}. % thesis - now split into sep table
%While obviously not as strong as full verification,
These completions guarantee that, if the test program could expose a bug,
it would depend on a data race not discovered yet,
or be reachable only under a superset combination of preemption points not yet reached.

@@

\begin{figure}[t]
	\begin{center}
		\includegraphics[width=0.88\textwidth]{nondets.pdf}
	\end{center}
	\caption[Nondeterministic data-race preemption points are required to find some bugs.]
	{Nondeterministic data-race preemption points are required to find some bugs.
	Requiring stateless model checking integration to identify to begin with,
	incorporating these races as new preemption points
	allowed Quicksand to find
	% lhb: 127.77%
	% phb: 191.30%
	128\% (Limited HB) to 191\% (Pure HB) as many data-race bugs
	compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}

>>>>

\begin{figure}[t]
	\begin{center}
		\begin{tabular}{p{0.91\textwidth}} % dont hyphenate 'overwrite'
			\begin{center}
			\begin{tabular}{rlcrl}
				\multicolumn{5}{c}{Initially {\tt int count = \const{0}; mutex\_t m;}} \\
				\\
				& {\bf Thread 1} & & & {\bf Thread 2} \\
				\hline
				11 & \texttt{\hilight{darkorange}{mutex\_lock}(\&m);} & &
						21 & \texttt{atomic\_xadd(\&count, \const{1});} \\
				12 & \texttt{count++;}                                & &
						22 & \texttt{\hilight{olivegreen}{yield}();} \\
				13 & \texttt{\hilight{darkblue}{mutex\_unlock}(\&m);} & &
						23 & \texttt{atomic\_xadd(\&count, \const{1});} \\
				14 & \texttt{assert(count >= \const{1});}             & &
						24 & \texttt{assert(count >= \const{2});} \\
			\end{tabular}
			\end{center}
			\\
			\begin{center}
			(a) Example program with synchronization API calls highlighted.
			\end{center}
			\\
			\begin{center}
			\begin{tabular}{llllc}
				& {\bf Thread 1} & & {\bf Thread 2} & {\bf Value of {\tt count}} \\
				\hline
				11 & \texttt{\hilight{darkorange}{mutex\_lock}(\&m);} & & & 0 \\
				12a & \texttt{int tmp = count;} & & & 0 \\
							& & 21 & \texttt{atomic\_xadd(\&count, \const{1});} & 1 \\
							& & 22 & \texttt{\hilight{olivegreen}{yield}();} & 1 \\
							& & 23 & \texttt{atomic\_xadd(\&count, \const{1});} & 2 \\
				12b & \texttt{count = tmp + 1;} & & & 1 \\
				13 & \texttt{\hilight{darkblue}{mutex\_unlock}(\&m);} & & & 1 \\
				14 & \texttt{assert(count >= \const{1});} & & & 1 \\
							& & 24 & \texttt{\hilight{assertfail}{assert(count >= 2);}} & 1 \\
			\end{tabular}
			\end{center}
			\\
			%(b) Interleaving of (a) in which Thread 2's assertion fails,
			%as the single increment of Thread 1 overwrites both those of Thread 2.
			(b) Buggy interleaving of (a) in which
			the single increment of Thread 1 overwrites both those of Thread 2.
			Note the data-race preemption at 12a-12b.
		\end{tabular}
	\end{center}
	\caption[Example bug requiring data-race preemption points to expose.]
	{Example bug requiring data-race preemption points to expose.
	Because the two threads use different modes of synchronization to protect their respective accesses to {\tt count},
	preempting on synchronization calls alone is insufficient to expose the bug.
	Rather, Thread 1 must be preempted between its non-atomic load and store of {\tt count}.
	}
	\label{fig:pps-example}
\end{figure}

@@

On 229 tests, SSS-MC-ICB timed out after 10 hours with no bugs found. % thesis note - pretty sure this *isn't* preempt-everywhere
Among these tests, QS-Limited-HB % thesis note -- i think that's what past-me meant. they just wrote, 'quicksand' but.
found bugs in 37.
The other 192 represent cases where neither Quicksand nor ICB were able to provide a conclusive result either way.%
\footnote{Thesis note: SSS-MC-Shared-Mem was added
%in a subsequent revision to \cite{quicksand}
%from when this analysis was conducted
subsequently to this analysis's publication in \cite{quicksand},
at which time SSS-MC-ICB was the best-performing approach among control experiments.
Nevertheless, mutual timeouts among all six testing approaches constituted roughly one third of the test suite.}
For these 192,
I show the number of state spaces Quicksand was able to complete in the
``Average tested subset SSes'' column.
% of \Cref{tab:drbugs}. % thesis - now split into sep table
%While obviously not as strong as full verification,
These completions guarantee that, if the test program could expose a bug,
it would depend on a data race not discovered yet,
or be reachable only under a superset combination of preemption points not yet reached.

@@

\begin{figure}[t]
	\begin{center}
		\includegraphics[width=0.87\textwidth]{nondets.pdf}
	\end{center}
	\caption[Nondeterministic data-race preemption points are required to find some bugs.]
	{Nondeterministic data-race preemption points are required to find some bugs.
	Requiring stateless model checking integration to identify to begin with,
	incorporating these races as new preemption points
	allowed Quicksand to find
	% lhb: 127.77%
	% phb: 191.30%
	128\% (Limited Happens-Before) to 191\% (Pure Happens-Before) as many data-race bugs
	compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}
